### 2012-08-29 sections/implementation.tex

\subsection{Splitter}
\subsection{Filter}
\subsection{Grouper}
\subsubsection{Group Aggregations}
\subsection{Group Filter}
\subsection{Merger}
\subsection{Ungrouper}