### 2012-08-31 sections/implementation.tex
* implementation   * flow query intermediate format     * custom c library to read flow records     * dedicated functions for each kind of operation   * execution workflow     * faster filter (with graph)     * grouper implementation       * possible grouping approaches       * faster grouping in generic case (nested bsearch)       * faster grouping in specific case (no bsearch needed) (with graph)       * cooked netflow v5 records for group aggregations     * group filter implementation     * merger implementation       * merger iterator utility       * skipping iterator permutations on sorte

\subsection{Flow Query Intermediate Format}
\input{sections/implementation/flow-query-intermediate-format}

\subsection{Execution Workflow}
\input{sections/implementation/execution-workflow}

\subsection{Adaptable Compression Levels}
\input{sections/implementation/adaptable-compression-levels}