### 2012-08-31 sections/implementation/adaptable-compression-levels.tex

The flow-records echoed to the standard output can also be written to a
Netflow $v5$ flow-tools file. The \texttt{--dirpath} switch allows one to
provide a directory path where the results can be stored.  Each stream is
stored as its own file with an ID to disambiguate it. Results from each stage
of the pipeline can also be written to separate files with the increase in the
verbosity level. In fact, \texttt{--dirpath} and \texttt{--verbose} work well
together to adjust the level at which the writes are to be made.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{zlevel}
    \caption{F$(v2)$: \texttt{z-level} Effect on Performance}
    \label{fig:engine-zlevel}
  \end{center}
\end{figure}

The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
the results written to the flow-tools files. \texttt{zlib} supports $9$
compression levels with $9$ being the highest. The \ac{NFQL} engine supports
\texttt{--zlevel} switch to allow the user to supply its desired choice of the
of the compression level. A default level of $5$ is used for writes if the
switch is not supplied during runtime.  Fig. \ref{fig:engine-zlevel} shows
the time taken to write a sample of records passing the filter stage for each
\texttt{z-level}. It goes to show that each level adds its own performance
overhead and must be used with discretion.