### 2012-08-31 sections/implementation/flow-query-intermediate-format.tex

A \ac{DNF} is a disjunction of conjunctive clauses. The elements of the
conjuctive clauses are terms. Each stage of the pipeline is represented in the
flowquery as a \ac{DNF} expression. The clauses in the \ac{DNF} are OR'd
together, while the terms in each clauses themselves are AND'd. The abstract
objects that store the JSON query and the results that incubate from each
stage are designed to be self-descriptive and hierarchically chainable. The
complete JSON query information for instance, is held within the
\texttt{flowquery} struct. Each individual branch of the flowquery itself is
described in a \texttt{branch} struct. A collection of these \texttt{branch}
structs are referenced in the parent \texttt{flowquery} struct.