### 2012-09-01 sections/conclusion.tex
* introduction   * traffic measurement approach blurb   * flow export protocol blurb * related work   * silk; and how it differs from nfql * performance evaluations   * stressing the filter   * stressing the grouper   * stressing the group filter   * stressing the merger   * stressing the ungrouper * conclusion

We presented \ac{NFQL}, an efficient C implementation of the stream-based flow
query language. The language allows applying absolute (or relative) filters,
aggregating flows into groups, evaluating intricate timing relationships among
them, and merging them into one collection. \ac{NFQL} can execute such complex
query tasks in matter of minutes, thereby expanding the scope of current flow
record processing tools. The conducted performance evaluations reveal that
\ac{NFQL} is on par with tools that support only absolute filters. SiLK, the
only package that provides tools that are similar to the rest of \ac{NFQL}'s
processing pipeline appears faster. This is because it can optimize its
operations in favor of the limited set of comparisons that are only based on
equality, and its usage of a proprietory file storage format.
\todo{...}