### 2012-09-01 sections/evaluation.tex

We used a public flow trace containing $20M$ records to run a number of flow
queries using our benchmarking suite. We used trace $7$
\footnote{\url{http://traces.simpleweb.org/traces/netflow/netflow1/netflow000.tar.bz2}},
from Simpleweb \footnote{Simpleweb is a data repository of traffic traces from
University of Twente}.  The input trace was compressed at \texttt{ZLIB\_LEVEL}
$5$ using the \texttt{zlib} suite. It was also converted to \texttt{nfdump}
and SiLK compatible formats and compressed with \texttt{zlib}
keeping the same compression level.  The suite was run on a high-end
machine\footnote{\url{crystal.eecs.jacobs-university.de}} with $24$ cores of
$2.5$ GHz clock speed and $18$ GiB of memory.

The first set of queries attempt to stress the filter stage.  We use varying
values on the packet field offset to determine the amount of flow records that
are passed by the filter. The resultant filtered records are written to
\texttt{flow-tools} compatible file format and compressed at
\texttt{ZLIB\_LEVEL} $5$ using \texttt{zlib} suite. The ratio of the number of
filtered records in the output trace to the number of the flow records in the
input trace is plotted against time. The evaluation results are shown in
Fig. \ref{fig:benchmarks-filter}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{filter}
    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
    \label{fig:benchmarks-filter}
  \end{center}
\end{figure}

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then again SilK and \texttt{nfdump} also use their own
proprietary format for trace files. As a result, the amount of data that needs
to read (or written) may be different to what it is for \ac{NFQL} and
\texttt{flowtools}. On the other hand, \texttt{nfdump} appears to be
significantly faster than the rest. This is because \texttt{nfdump} lacks
\texttt{zlib} support, and as such the files that are read and written used
\texttt{lzo} compression scheme which trades space for achieving faster
compression and decompression. It is important to note, that all the tools
were single-threaded in this evaluation, and did not completely utilize the
$24$ cores that were at their disposal. It comes as a realization, that
filtering the input using multiple threads by memory mapping the trace and
adding \texttt{lzo} compression will drastically improve \ac{NFQL}'s filter
performance.

The second set of queries attempt to stress the grouper stage. We reuse the
filter query that produces a $1.0$ ratio to allow the grouper to receive the
entire trace as a filtered recordset. The grouper part of the query then
gradually increases the number of grouping terms in the \ac{DNF} expression to
increase the output/input ratio. The resultant groups are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of groups formed to the number of the input filtered
records is plotted against time. \texttt{nfdump} and \texttt{flowtools} do not
support grouping, and therefore are not considered in this evaluation. The
evaluation results are shown in Fig. \ref{fig:benchmarks-grouper}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{grouper}
    \caption{Grouper Stage: NFQL vs SiLK}
    \label{fig:benchmarks-grouper}
  \end{center}
\end{figure}

The evaluation graph reveals that the performance of the \ac{NFQL} grouper
stage is close. The time taken by the tools are comparable on lower ratios,
but on higher ratios, \ac{NFQL} starts to drift apart. Since most of the time
is taken in writing the records to files, it is unclear whether SiLK's usage
of a proprietary format which may reduce reads/writes is responsible for the
drift on higher ratios. SiLK's \texttt{rwgroup} tool is also supplied a
\texttt{--summarize} flag in all the queries. This gives SiLK the leverage to
not store information about which members are part of the group.  \ac{NFQL} on
the other hand needs to allocate resources (which may take time) to keep this
information in its data structures, since the ungrouper later may request to
write the members of a group while unfolding the tuples. It is also important
to note that both the tools again remained single-threaded throughout the
evaluation. SiLK took an advantage of an inherent concurency arising from how
the query is structured as one single \texttt{bash} script using pipes. The
pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two process run
concurrently, the effect of which gets more pronounced on higher ratios and
can be a drift determining factor.  The profiling results from GNU
\texttt{gprof} \cite{graham:1982} indicate that $60\%$ of the time is taken in
\texttt{qsort} comparator calls.  As a result, it comes as no surprise, that
bifurcating \texttt{qsort} invocation to multiple threads and later merging
the results back using merge sort will help parallelize the grouper stage and
maybe reduce the drift on higher ratios. In addition, since all of the
evaluation queries had grouping terms using an equality comparator, \ac{NFQL}
can introspect such a grouping rule to dynamically optimize processing
searches using a hashtable and turn to \texttt{qsort} based grouping only as a
fallback.

The third set of queries attempt to stress the group filter stage. We reuse
the filter and grouper queries that produce a $1.0$ ratio to allow the group
filter to receive the entire trace as input. This means, each flow record of
the original trace now becomes a group record for the group filter.  The group
filter then reuses the same varying values of the packet offset to determine
the amount of groups that are filtered ahead. The filtered groups are again
written as \texttt{flowtools} files using the same \texttt{zlib} compression
level. The ratio of the number of filtered groups formed to the number of the
input group records is plotted against time. The evaluation results are shown
in Fig. \ref{fig:benchmarks-groupfilter}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{groupfilter}
    \caption{Group Filter Stage: NFQL vs SiLK}
    \label{fig:benchmarks-groupfilter}
  \end{center}
\end{figure}

It can be seen that the timings of \ac{NFQL} are far apart from that of SiLK.
It is due to the drift already created by the grouper at $1.0$ ratio in the
previous stage. As a result, the group filter comes into play only after $300$
seconds, whereas SiLK's group filtering already starts just below $150$
seconds. Even if we normalize the graph, it can be observed that the group
filter has a significantly higher slope. This is because it is only executed
once the grouper returns, and therefore has to reiterate the groups to make a
filtering decision.

The fourth set of queries attempt to stress the merger stage. We reuse the
filter, grouper and group filter queries that produce a $1.0$ ratio. These
queries are then run in two separate branches to produce identical filtered
group records.  The merger then applies match rules to produce different
output to input ratios. The groups that are merged are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of merged groups to twice\footnote{Each branch pushes the
entire trace as an input to the merger.} the number of flow records in the
original trace is plotted against time. A data point for SiLK for the $0.2$
ratio is not available since the \ac{NFQL} query executed at that data point
uses \texttt{OR} expressions which are not supported by SiLK. As a result, an
equivalent SiLK query is not formulated.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{merger}
    \caption{Merger Stage}
    \label{fig:benchmarks-merger}
  \end{center}
\end{figure}

It can be seen that the merger is the most performance hit and time consuming
stage of the \ac{NFQL} pipeline thus far. It is due to the fact that the
merger is working on twice the number of flow records than any other previous
stage. In addition, each branch is writing the results of the filter, grouper
and group filter stage to \texttt{flowtools} files. As a result, the amount of
disk I/O involved is twice as much as well. Even though each branch is
delegated to a separate core using affinity masks, most of time is taken in
writing these results to the file. These results although look less promising,
they are way better than the previous merger implementation. The newer merger
takes advantage of sorted nature of filtered groups and therefore can
signifantly reduce the number of merger matches. It also writes a merged group
record to file only once despite the number of times it has matched. Without
these optimizations, running such queries on the merger would keep the CPU
churning for days without results.

The last set of queries attempt to stress the ungrouper stage. They reuse the
entire merger queries as is, but enable the ungrouper now as well. This means,
that the ungrouper now attempts to unfold the merged groups returned by the
merger to write their member flow records to \texttt{flowtools} files.
However, since the merger receives each flow record as its own filtered group,
each merged group has only one member. As a result, the ungrouper ends up
executing its logic and rewriting the merged groups as \texttt{flowtools}
files using the using the same \texttt{zlib} compression level. The ratio of
the number of result flow records to twice the number of the flow records in
the input trace is plotted against time. The evaluation results are shown in
Fig. \ref{fig:benchmarks-ungrouper}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{ungrouper}
    \caption{Ungrouper Stage}
    \label{fig:benchmarks-ungrouper}
  \end{center}
\end{figure}

It can be seen that the evaluation behavior is similar to that shown by the
merger, but is just more involved. This is because the ungrouper also has to
write the entire merged groupset to file, as is done by merger. This is the
reason for the execution engine in taking twice the amount of time. In fact,
this evaluation is performed only to stress the functioning of the ungrouper
and running queries that merge each filtered group filtered individually is
less useful in practise. Ideally, the execution engine must shutdown the
ungrouper if the merger merges every filtered group record in its output. This
is because, such a behavior implicitly eliminates the need of the ungrouper.