### 2012-09-02 sections/implementation.tex

\subsection{Flow Query Intermediate Format}
\input{sections/implementation/flow-query-intermediate-format}
\label{subsec:intermediate-format}

\subsection{Execution Workflow}
\input{sections/implementation/execution-workflow}
\label{subsec:execution-workflow}

\subsection{Adaptable Compression Levels}
\input{sections/implementation/adaptable-compression-levels}
\label{subsec:adaptable-compression-levels}

\subsection{Performance Optimizations}
\input{sections/implementation/performance-optimizations}
\label{subsec:performance-optimizations}