### 2012-09-02 sections/conclusion.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

We presented \ac{NFQL}, an efficient C implementation of the stream-based flow
query language. The language allows applying absolute (or relative) filters,
aggregating flows into groups, evaluating intricate timing relationships among
them, and merging them into one collection. \ac{NFQL} can execute such complex
query tasks in matter of minutes, thereby expanding the scope of current flow
record processing tools. The conducted performance evaluations reveal that
\ac{NFQL} is on par with tools that support only absolute filters. SiLK, the
only openly available package that provides tools that are similar to the rest
of \ac{NFQL}'s processing pipeline appears faster. This is because it can
optimize its operations in favor of the limited set of comparisons that are
only based on equality, and its usage of a different file storage format.
\todo{...}