### 2012-09-02 sections/implementation/execution-workflow.tex

A custom C library has been written to directly read/write data in the
\texttt{flow-tools} format. The library sequentially reads the complete
flow-records into memory to support random access required for relative
filtering. Each flow-record is stored in a \texttt{char} array and the offsets
to each field are stored in separate \texttt{structs} as shown in listing
\ref{lst:c-lib-struct}. The array of such records are indexed allowing fast
retrieval in $O(1)$ time.

\lstset{caption=Trace Data Struct,
				tabsize=2, language=C, numbers=left,stepnumber=1,
        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
        keywordstyle=\color{blue}, xleftmargin=15pt,
        rulesepcolor=\color{black}, label=lst:c-lib-struct,
        aboveskip=10pt, captionpos=b}
\begin{lstlisting}
struct ft_data {
  int                                         fd;
  struct ftio                                 io;
  struct fts3rec_offsets                      offsets;
  struct ftver                                version;
  u_int64_t                                   xfield;
  int                                         rec_size;
  char**                                      recordset;
  size_t                                      num_records;
};
\end{lstlisting}

A sample JSON query \ac{DNF} term definition is shown in listing
\ref{lst:filter-struct}. In order to be able to make comparisons on the field
offsets of such a term, the comparator needs to know the type of the
comparison and the length of the field offset. This information is parsed by
execution engine once the query is read and is therefore not available at
compile time.  In order to subvert the need to define complex branching
statements, a dedicated comparator is defined for every possible field length
and comparison operation. These comparators are generated using a Python
script. This allows the term definitions to make runtime calls using a
function name derived from the combination of operation type and field length.

\lstset{caption=Filter Term Struct,
				tabsize=2, language=C, numbers=left,stepnumber=1,
        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
        keywordstyle=\color{blue}, xleftmargin=15pt,
        rulesepcolor=\color{black}, label=lst:filter-struct,
        aboveskip=10pt, captionpos=b}
\begin{lstlisting}
struct filter_term {
  size_t                                      field_offset;
  uint64_t                                    value;
  uint64_t                                    delta;
  struct filter_op*                           op;
  bool (*func)(
               const char* const              record,
               size_t                         field_offset,
               uint64_t                       value,
               uint64_t                       delta
              );
};
\end{lstlisting}

\subsubsection{Splitter} \ac{NFQL} uses identifiers to reference a flow record
in the \texttt{char} array. This gives away the need to every time copy all
the flow-records when moving ahead in the processing pipeline. As a result,
there is no dedicated splitter stage in the execution engine. Each branch
references the flow records from a common memory location. This helps keep the
memory costs at a minimum when multiple branches are involved.

\subsubsection{Filter}
\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=1.0\linewidth]{filter-fv1-fv2}
    \caption{Filter Stage: NFQL vs NFQL (optimized)}
    \label{fig:fv1-fv2-filter}
  \end{center}
\end{figure}

The execution engine, as defined by the flow query language must read all the
flow records of a supplied trace into memory before starting the processing
pipeline.  Since, the filter stage uses the supplied set of absolute rules to
make a decision on whether or not to keep a flow record; it has to pass
through the whole in-memory recordset \emph{again} to fill in the filter
results. This technique involves multiple linear runs on the trace and
therefore slows down when the ratio of number of filtered records to the total
number of flow-records is high. We optimized this behavior in \ac{NFQL} by
merging the filter stage with in-memory read of the trace. This means, a
decision on whether or not to make room for a record in memory and eventually
hold a pointer for it in filter results is done upfront as soon as the record
is read from the trace. In addition, if a request to write the filter stage
results to a flow-tools file has been made, the writes are also made as soon
the filter stage decision is available, thereby allowing
reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
records in the trace. We used the publically available Trace 7 from the
SimpleWeb \cite{simpleweb} to compare the performance of the optimized
\ac{NFQL} implementation against the one defined by the flow query language as
shown in Fig. \ref{fig:fv1-fv2-filter}. The filter stage implementation with
these optimizations runs 10 times faster and is more pronounced on higher
ratios.

\subsubsection{Grouper}

In order to be able to make relative comparisons on field offsets, a simple
approach is to linearly walk through each filtered record against the filtered
recordset leading to a complexity of $O(n^2)$, where $n$ is the number of
filtered records. A smarter approach is to put the copy in a hash table and
then try to map each pointer while walking down the filtered recordset once,
leading to a complexity of $O(n)$. The hash table approach however, will only
work on equality comparisons. A better approach, as implemented in \ac{NFQL}
is to sort the filtered recordset on all the requested grouping terms of a
clause. This helps the execution engine perform a nested binary search to
reduce the linear pass to a fairly small filtered recordset. However, the
number of elements upon which the search has to be performed needs to be known
at each level of the binary lookup. In order to avoid a linear run to count
the number of elements, the parent level binary search invocation returns a
boundary with the first and last element to enable such a calculation. This
helps the grouper perform faster search lookups to find records that must
group together in $O(n*lg(k))$ time with a preprocessing step taking
$O(p*n*lg(n))$ in the average case, where $n$ is the number of filtered
records, $p$ is the number of grouping terms in a clause and $k$ is the number
of unique filtered records.

\begin{figure}[h!]
  \begin{center}
    \includegraphics*[width=1.0\linewidth]{grouper-fv2-bsearch-fv2}
    \caption{Grouper Stage: NFQL (Generic) vs NFQL (Specific)}
    \label{fig:fv1-fv2-grouper}
  \end{center}
\end{figure}

The grouping approach has further been optimized when the filtered records are
grouped for equality. In such a scenario, the need to search for unique
records and a  subsequent binary search goes away.  The groups can now be
formed in $O(n)$ time with the same preprocessing step taking $O(p*n*lg(n)$.
The performance evaluation of the grouper handling this special case against
its behavior when handling generic cases is shown in Fig.
\ref{fig:fv1-fv2-grouper}. Profiling the performance of the grouper on queries
that produce higher ratios reveal that a significant amount of time is taken
in binary search which is eliminated in the special case of equality
comparisons.

The resultant group records are a conglomeration of multiple flow records with
some common characteristics. Some of the non-common characteristics can also
be aggregated into a single value using group aggregations as defined in the
query. Such an aggregated group record is again mapped to a NetFlow $v5$
record template. This allows the aggregated group records to be written to
file as a representative of all its members.

\subsubsection{Group Filter} The groupfilter stage is used to filter the
groups produced by the grouper based on some absolute rules defined in a
\ac{DNF} expression. The \texttt{struct} term holds information about the flow
record offset, the value being compared to and the operator which maps to a
unique \texttt{enum} value.  This enum value is used to map the operation to a
specific group-filter function. The group-filter functions are auto-generated
using a Python script.

\subsubsection{Merger}

The merger is used to relate groups from different branches according to a
merging criterean. The implementation is not trivial since the number of
branches that need to be spawned is read from the query and is not known until
runtime. As a result, an iterator that can provide all possible permutations
of $m-$tuple (where $m$ is the number of branches) group record $ID$s was
needed.  The result of the iterator can then be used to make a match. The
merger stage, begins by initializing this iterator passing it the number of
branches, and information about each branch. Then, it loops over to get a new
$m-$tuple of group record $ID$s on each iteration until the iterator returns
\texttt{false}.

The merger as formulated in the flow query language needs to match each group
record from one branch with every other record of each branch. This leads to a
complexity of $O(g^m)$ where $g$ is the number of filtered group records and
$m$ is the number of branches. The possible number of tries when matching
group records however can be reduced by sorting the group records on the field
offsets used for a match. The \ac{NFQL} implementation optimizes the merger to
skip over iterator permutations when a state of a current field offset value
may not allow any further match beyond the index in the current branch.  For
such an optimization to work, the filtered group records must be sorted in the
order of field offsets specified in the merger clause. Specifying the filtered
group records in any other order may lead to undefined behavior. This means,
that if the same field offsets were used in the grouper stage, the terms in
the group clause can be rearranged by the query designer to align with the
order of terms in the merger clause.

The flow query language also bases the merger matches on the notion of matched
tuples. This means that a filtered group record can be written to a file
multiple times if it is part of multiple matched tuples. This situation is
very common and it worsens when different branches have similar filtered
groups records. Since, the function of the merger is to find a match of groups
records across branches based on a predefined condition, all the group records
across branches that satisfy the condition can be clubbed into one collection
instead of separate tuples. All the group records within a collection can then
be written to the file. This eliminates the inherent redundancy and
significantly improves the merger performance.

The performance comparison of this approach against the one suggested by the
specification is tricky. The merger implementation of the original
specification is slow. It is so slow that it keeps churning the CPU for days
without results. The newer approach takes less than an hour. The performance
evaluation of this newer approach is discussed in more details in the next
section

\subsubsection{Ungrouper} The approach of clubbing the merged group records into
a collection incurs a reimplementation of the ungrouper. The ungrouper, as a
result accepts a collection of matched filtered group records as input. It
then iterates over each collection to unfold it groups and write their flow
record members to files.