### 2012-09-04 sections/implementation/flow-query-intermediate-format.tex
* revision with professor's comments   * introduction     * updated title, added vlad's IM 2011 paper     * reduced from bibtex entries, back to page 8     * added collado:2009 and sperotto:2010 references   * implementation     * flow query intermediate format     * grouper, grouper aggregation     * merger (TODO)

Each stage of the processing pipeline is expressed in the \texttt{JSON} query
as a \ac{DNF} expression. A \ac{DNF} expression is a disjunction of
conjunctive clauses. The elements of the conjuctive clauses are terms. The
clauses in the \ac{DNF} are OR'd together, while the terms in each clause
themselves are AND'd. \texttt{json-c} \cite{jsonc} is used to parse the flow
query file. The pipeline stages of the flow query language are also
encapsulated in Python classes and scripts have been written to facilitate the
front-end parser to serialize these pipeline objects to a \texttt{JSON} flow
query.  The mapping of the query to the structs defined in the execution
engine is not trivial. When reading the \texttt{JSON} query at runtime, the
field names of a NetFlow $v5$ record are read in as strings.  Utility
functions are defined that map the field names to internal struct offsets and
the field types and the operations to internal unique enum members.

The abstract objects that store the JSON query and the results that incubate
from each stage are designed to be self-descriptive and hierarchically
chainable.  The complete JSON query information for instance, is held within
the \texttt{flowquery} struct as shown in listing \ref{lst:struct}. Each
individual branch of the flowquery itself is described in a \texttt{branch}
struct.

\lstset{caption=Flow Query and Branch Structs,
				tabsize=2, language=C, numbers=left,stepnumber=1,
        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
        keywordstyle=\color{blue}, xleftmargin=15pt,
        rulesepcolor=\color{black}, label=lst:struct,
        aboveskip=10pt, captionpos=b}
\begin{lstlisting}
struct flowquery {
  size_t                                      num_branches;
  size_t                                      num_merger_clauses;

  struct branch**                             branchset;
  struct merger_clause**                      merger_clauseset;
  struct merger_result*                       merger_result;
  struct ungrouper_result*                    ungrouper_result;
};

struct branch {
  int                                         branch_id;
  struct ftio*                                ftio_out;
  struct ft_data*                             data;

  size_t                                      num_filter_clauses;
  size_t                                      num_grouper_clauses;
  size_t                                      num_aggr_clause_terms;
  size_t                                      num_groupfilter_clauses;

  struct filter_clause**                      filter_clauseset;
  struct grouper_clause**                     grouper_clauseset;
  struct aggr_term**                          aggr_clause_termset;
  struct groupfilter_clause**                 groupfilter_clauseset;

  struct filter_result*                       filter_result;
  struct grouper_result*                      grouper_result;
  struct groupfilter_result*                  gfilter_result;
};
\end{lstlisting}

The \texttt{JSON} query can also disable the stages at runtime.  This means
that one only has to supply the constructs that one wishes to use.  The
constructs that are not defined in the \texttt{JSON} query are inferred by the
execution engine as a disable request. The execution engine uses disable flags
that are turned on when the query is parsed. These flags are used throughout
the engine to only enable the requested functionality.