### 2012-09-04 sections/introduction.tex

Researchers, service providers and security analysts have long been interested
in network and user behavioral patterns of the traffic crossing the internet
backbone. They want to use this information for the purpose of billing and
mediation, bandwidth provisioning, detecting malicious attacks, network
performance evaluation and overall improvement. Traffic measurement techniques
that have been rapidly evolving in the last decade, have matured enough today
to provide such an insight.

Capturing network flow records has emerged out to be one of the favored
network measurement techniques. In this technique, packets traversing an
obvervation point are not captured raw, instead they are aggregated together
based on some common characteristics. The common characteristics are learnt by
inspecting the packet headers as they cross the observation point.
Flow-records resulting from such an aggregation are then exported to a
collector for further analysis. This reduces the amount of traffic aggregated
at the collector.

NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
export. NetFlow \cite{rfc3954} is a proprietary network protocol designed by
Cisco Systems. It allows routers to generate and export flow records to a
designated collector. NetFlow v$9$ provides flexibility of user-tailored
export templates, \ac{MPLS} and IPv$6$ support and a larger set of flow keys.
\ac{IPFIX} \cite{rfc5101} on the other hand is an open standard by IETF and
deemed to be the logical successor of NetFlow v$9$ on which it is based. The
novelty of the standard lies in its ability to describe record formats at
runtime using templates based on an extensible and well-defined information
model. The data transfer mechanism is also simplistic and extensible by being
unidirectional and transport protocol agnostic.

The wide applicability of this approach is easily seen from the pervasive use
of flow records for a vibrant set of network analysis applications. For
instance, the authors in \cite{mkim:2004} use the flow characteristics in the
traffic pattern to formalize a detection function that maps traffic patterns
to different DoS attacks, whereas in \cite{sdominik:2010}, the authors
use the flow-record data to exploit timing characteristics of webmail clients
to classify features that could identify webmail traffic from any other
traffic running over HTTPS.

Understanding intricate traffic patterns require sophisticated flow analysis
tools that can mine flow records for such a usage.  Unfortunately current
tools fail to deliver owing to their language design and simplistic filtering
methods.  We recently proposed a flow query language design
\cite{vmarinov:2009} that aims to cater to such needs.  It can process
flow records, aggregate them into groups, apply absolute (or relative) filters
and invoke Allen interval algebra rules \cite{fallen:1983}. The applicability
of the language can be seen from \cite{vperelman:2011} where the authors
formulate flow queries to identify flow signatures of popular applications.

In this paper, we introduce \ac{NFQL}, an efficient C implementation of the
flow query language. It is the next iteration of the first prototype
implementation, Flowy \cite{kkanev:2010}, which was written in Python.
\ac{NFQL}, however, is not just a reimplementation in a new language, but the
inner workings have been reimagined to allow the execution engine to scale to
real-world sized traces. This has been possible by replacing the performance
critical stages of the pipeline with crispier algorithms that lower the
execution times, making \ac{NFQL} comparable to contemporary flow analysis
tools.

The paper is organized as follows. In section \ref{sec:relatedwork} we survey
the current state-of-the-art flow-processing tools and reason how \ac{NFQL} is
different from each one of them. In section \ref{sec:design} we describe the
flow query language by discussing each stage of the processing pipeline. In
section \ref{sec:implementation} we introduce the inner workings of \ac{NFQL}.
It begins by providing an overview of the \ac{NFQL} architecture and the
structure of the intermediate format used to exchange messages. The workflow
of the execution engine is described next and is supplemented by a number of
performance optimizations made to make the implementation scale. The section
is followed by performance evaluations comparing \ac{NFQL} against
contemporary flow processing tools alongwith with a discussion on its current
limitation and future outlook in section \ref{sec:evaluation}, with conclusion
in section \ref{sec:conclusion}.

\begin{figure*}[!t]
\centering
\includegraphics[width=0.9\linewidth]{nfql-pipeline}
\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
\label{fig:nfql-pipeline}
\end{figure*}