### 2012-09-04 sections/implementation/adaptable-compression-levels.tex
* revision with professor's comments   * performance optimizations

The flow-records echoed to the standard output can also be written to Netflow
$v5$ \texttt{flow-tools} files. The \texttt{-{}-dirpath} switch allows one to
provide a directory path where the results can be stored. Results from each
stage of the pipeline can also be written to separate files with the increase
in the verbosity level. In fact, \texttt{-{}-dirpath} and \texttt{-{}-verbose}
work well together to adjust the level at which the writes are to be made.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{zlevel}
    \caption{\texttt{z-level} Effect on Performance}
    \label{fig:engine-zlevel}
  \end{center}
\end{figure}

The engine uses the \texttt{zlib} \cite{rfc1950} software library to compress
the results written to the \texttt{flow-tools} files. \texttt{zlib} supports
$9$ compression levels with $9$ being the highest. The \ac{NFQL} engine
supports \texttt{-{}-zlevel} switch to allow the user to supply its desired
choice of the of the compression level. A default level of $5$ is used for
writes if the switch is not supplied during runtime.  Fig.
\ref{fig:engine-zlevel} shows the time taken to write a sample of records
passing the filter stage for each \texttt{z-level}. It goes to show that each
level adds its own performance overhead and must be used with discretion. It
is also important to note that other tools may use different compression
algorithms. \texttt{nfdump} for instance, uses \texttt{lzo} \cite{lzo}
compression to trade space for faster compression and decompression.