### 2012-09-04 sections/evaluation.tex
* revision with professor's comments   * performance evaluations

We used the first $20M$ records from a public flow trace to run a number of
flow queries using our benchmarking suite. We used the publically available
Trace 7 from the SimpleWeb \cite{simpleweb} repository for all the performance
evaluations. The input trace was compressed at \texttt{ZLIB\_LEVEL} $5$ using
the \texttt{zlib} suite. It was also converted to \texttt{nfdump} and SiLK
compatible formats and compressed with \texttt{zlib} keeping the same
compression level. The suite was run on a machine with $24$ cores of
$2.5$ GHz clock speed and $18$ GiB of memory.

The first set of queries attempt to stress the filter stage.  We use varying
values on the \texttt{packet} field offset to determine the amount of flow
records that are passed by the filter. The resultant filtered records are
written to \texttt{flow-tools} compatible file format and compressed at
\texttt{ZLIB\_LEVEL} $5$. The ratio of the number of filtered records in the
output trace to the number of the flow records in the input trace is plotted
against time. The evaluation results are shown in Fig.
\ref{fig:benchmarks-filter}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{filter}
    \caption{Filter Stage: NFQL vs SiLK, Flow-Tools, Nfdump}
    \label{fig:benchmarks-filter}
  \end{center}
\end{figure}

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then again SilK and \texttt{nfdump} also use their own file
format. As a result, the amount of data that needs to be read (or written) may
be different to what it is for \ac{NFQL} and \texttt{flowtools}.
\texttt{nfdump} appears to be significantly faster than the rest. This is
because \texttt{nfdump} lacks \texttt{zlib} support, and as such the files
that are read and written use the \texttt{lzo} compression scheme. It is
important to note, that all the tools were single-threaded in this evaluation,
and did not completely utilize the $24$ available cores. It comes as a
realization, that adding \texttt{lzo} compression will drastically improve
\ac{NFQL}'s filter performance.

The second set of queries attempt to stress the grouper stage. We reuse the
filter query that produces a $1.0$ ratio to allow the grouper to receive the
entire trace as a filtered recordset. The grouper part of the query then
gradually increases the number of grouping terms in the \ac{DNF} expression to
increase the output/input ratio. The resultant groups are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of groups formed to the number of the input filtered
records is plotted against time. \texttt{nfdump} and \texttt{flowtools} do not
support grouping, and therefore are not considered in this evaluation. The
results are shown in Fig. \ref{fig:benchmarks-grouper}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{grouper}
    \caption{Grouper Stage: NFQL vs SiLK}
    \label{fig:benchmarks-grouper}
  \end{center}
\end{figure}

The evaluation graph reveals that the performance of the \ac{NFQL} grouper
stage is close to the performance of SiLK. The time taken by the tools are
comparable on lower ratios, but on higher ratios, \ac{NFQL} starts to drift
apart. Since most of the time is taken in writing the records to files, it is
unclear whether SiLK's usage of its own file format, which may reduce
reads/writes, is responsible for the drift on higher ratios. SiLK's
\texttt{rwgroup} tool is also supplied a \texttt{-{}-summarize} flag to force it
to write only the first record of each group, to make both tools write the
same number of records to files. This gives SiLK the leverage to not store
information about which members are part of the group. \ac{NFQL} on the other
hand needs to allocate resources (which may take time) to keep this
information in its data structures, since the ungrouper later may request to
write the members of a group while unfolding the tuples. It is also important
to note that both the tools again remained single-threaded throughout the
evaluation. SiLK took an advantage of an inherent concurency arising from how
the query is structured as one single \texttt{bash} script using pipes. The
pipe between \texttt{rwsort} and \texttt{rwgroup} makes the two processes run
concurrently, the effect of which gets more pronounced on higher ratios and
can be a drift determining factor.  The profiling results from GNU
\texttt{gprof} \cite{graham:1982} indicate that $60\%$ of the time is taken in
\texttt{qsort} comparator calls.  As a result, it comes as no surprise, that
bifurcating \texttt{qsort} invocation to multiple threads and later merging
the results back using merge sort will help parallelize the grouper stage and
maybe reduce the drift on higher ratios. In addition, since all of the
evaluation queries had grouping terms using an equality comparator, \ac{NFQL}
can introspect such a grouping rule to dynamically optimize processing
searches using a hashtable and turn to \texttt{qsort} based grouping only as a
fallback.

The third set of queries attempt to stress the group filter stage. We reuse
the filter and grouper queries that produce a $1.0$ ratio to allow the group
filter to receive the entire trace as input. This means, each flow record of
the original trace now becomes a group record for the group filter.  The group
filter then reuses the same varying values of the \texttt{packet} field offset
to determine the amount of groups that are filtered ahead. The filtered groups
are again written as \texttt{flowtools} files using the same \texttt{zlib}
compression level. The ratio of the number of output filtered groups to the
number of the input group is plotted against time. The results are shown in
Fig. \ref{fig:benchmarks-groupfilter}.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{groupfilter}
    \caption{Group Filter Stage: NFQL vs SiLK}
    \label{fig:benchmarks-groupfilter}
  \end{center}
\end{figure}

It can be seen that the timings of \ac{NFQL} are far apart from that of SiLK.
It is due to the drift already created by the grouper at the $1.0$ ratio in
the previous stage. As a result, the group filter comes into play only after
$300$ seconds, whereas SiLK's group filtering already starts just below $150$
seconds. Even if we normalize the graph, it can be observed that the \ac{NFQL}
group filter has a higher slope. This is because it is only executed once the
grouper returns, and therefore has to reiterate the groups to make a filtering
decision.

The fourth set of queries attempt to stress the merger stage. We reuse the
filter, grouper and group filter queries that produce a $1.0$ ratio. These
queries are then run in two separate branches to produce identical filtered
group records.  The merger then applies match rules to produce different
output to input ratios. The groups that are merged are again written as
\texttt{flowtools} files using the same \texttt{zlib} compression level. The
ratio of the number of merged groups to twice\footnote{Each branch pushes the
entire trace as an input to the merger.} the number of flow records in the
original trace is plotted against time. A data point for SiLK for the $0.2$
ratio is not available since the \ac{NFQL} query executed at that data point
uses \texttt{OR} expressions, which are not supported by SiLK. As a result, an
equivalent SiLK query is not formulated.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{merger}
    \caption{Merger Stage}
    \label{fig:benchmarks-merger}
  \end{center}
\end{figure}

It can be seen that the merger is the most performance critical stage of the
\ac{NFQL} pipeline thus far. It is due to the fact that the merger is working
on twice the number of flow records than any other previous stage. In
addition, each branch is writing the results of the filter, grouper and group
filter stage to \texttt{flowtools} files. As a result, the amount of disk I/O
involved is twice as much as well. Even though each branch is delegated to a
separate core using affinity masks, most of time is taken in writing these
results to the file. These results although look less promising, they are way
better than the original \ac{NFQL} merger implementation. The optimized merger
implementation takes advantage of sorted nature of filtered groups and
therefore can significantly reduce the number of merger matches. It also writes
a merged group record to a file only once despite the number of times it has
matched. Without these optimizations, running such queries on the merger would
keep the CPU churning for days without results.

The last set of queries attempt to stress the ungrouper stage. They reuse the
entire merger queries as is, but enable the ungrouper as well. This means,
that the ungrouper now attempts to unfold the merged groups returned by the
merger to write their member flow records to \texttt{flowtools} files.
However, since the merger receives each flow record as its own filtered group,
each merged group has only one member. As a result, the ungrouper ends up
rewriting the merged groups as \texttt{flowtools} files using the same
\texttt{zlib} compression level. The ratio of the number of result flow
records to twice the number of the flow records in the input trace is plotted
against time. The evaluation results are shown in Fig.
\ref{fig:benchmarks-ungrouper}.

\begin{figure}[h!]
  \begin{center}
    \includegraphics* [width=0.9\linewidth]{ungrouper}
    \caption{Ungrouper Stage}
    \label{fig:benchmarks-ungrouper}
  \end{center}
\end{figure}

It can be seen that the behavior of the ungrouper is similar to that shown by
the merger. Since, the ungrouper also has to write the entire merged groupset
to files, as is previously done by merger, the execution engine takes twice the
amount of time. It is important to note, that SiLK does not have such an
equivalent ungrouping tool and is therefore not considered in this final
evaluation.