### 2012-09-04 sections/conclusion.tex
revision with professor's comments   * conclusion, with blurb on usefulness of benchmarking queries   * removed ungrouper graph

We presented \ac{NFQL}, an efficient C implementation of the stream-based flow
query language. The language allows applying absolute (or relative) filters,
aggregating flows into groups, evaluating timing relationships among them, and
merging them into one collection. \ac{NFQL} can execute such complex queries
in matter of minutes, thereby expanding the scope of current flow record
processing tools. The conducted performance evaluations reveal that \ac{NFQL}
is on par with tools that support only absolute filters. SiLK, the only openly
available package that provides tools that are similar to the rest of
\ac{NFQL}'s processing pipeline appears faster. This is because it can
optimize its operations in favor of the limited set of comparisons that are
only based on equality, and its usage of a different file storage format. The
evaluation queries developed as a part of this research work may also develop
into a more general benchmark suite for flow query tools and platforms.