### 2012-09-04 sections/design.tex
* revision with professor's comments   * flow query language design

The flow query language consists of a number of independent stages that are
connected to one another to form a processing pipeline.  Each stage receives
the content from the previous, performs an operation and pushes the results to
the next stage as shown in Fig. \ref{fig:nfql-pipeline}. A complete
description of the semantics of each pipeline stage can be found in
\cite{vmarinov:2009}.

The pipeline starts with the splitter. It takes the flow-records data as input
in \texttt{flow-tools} compatible format. The splitter is responsible to
duplicate the input data out to several branches without any processing
whatsoever. This allows each of the branches to receive an identical copy of
the flow data to process it independently.

The filter is the first processing stage of a branch in the pipeline.  It
performs \emph{absolute} filtering on the input flow-records data.  The
flow-records that pass the filtering criterion are forwarded to the grouper,
the rest of the flow-records are dropped. The filter compares separate fields
of a flow-record against either a constant value or a value on a different
field of the \emph{same} flow-record. The filter cannot \emph{relatively}
compare two different incoming flow-records.

The grouper performs aggregation of the input flow-records data. It consists
of a number of terms that correspond to a specific clause. A flow-record in
order to be a part of the group should satisfy at-least one clause. A
flow-record, although, can be a part of multiple clauses, it cannot be part of
multiple groups. The grouping terms can be either absolute or relative. The
newly formed groups, which are passed on to the group filter, can also contain
meta-information about the flow-records contained within the group using the
aggregate clause defined as part of the grouper query.

The group-filter is the last processing stage of a branch. It performs
\emph{absolute} filtering on group-records. The group-records
that pass the filtering criterion are forwarded to the merger, the rest of the
group-records are dropped. The group-filter compares separate fields (or
aggregated fields) of a group-record against either a constant value or a value
on a different field of the \emph{same} group-record. The group-filter cannot
\emph{relatively} compare two different incoming group-records.

The merger performs relative filtering on the N-tuples of groups formed from
the N stream of groups passed on from the group-filter as input. The relative
filtering on the groups are applied to express timing and concurrency
constraints using Allen interval algebra \cite{fallen:1983}. The ungrouper is
the final processing element of the pipeline. It unwraps the tuples of
group-records into individual flow-records, ordered by their timestamps. Any
duplicate flow-records appearing from several group-records are eliminated and
are sent as output only once.