### 2012-10-15 rebuttal.txt
make some initial changes to rebuttal.txt

Review 1:
1. wrong track
2. lack of description of the language
3. too many implementation details
4. metrics used
5. research questions and conclusions

Review 2:
1. Gigascope, point to the original NFQL paper

Review 3:
1. Explain that NFQL is defined elsewhere, this paper is discussing
   implementation aspects.
2. Discuss storage requirement vs. state requirement (after all NFQL
   is a streaming language but it requires state to be maintained)

Review 4:
1. Explain what NFQL can do that other tools can't do
2. Clarify the order splitter/filter in the query language and the
   order of filter/splitter in the implementation

JoSch's notes/comments
======================

Review 1
--------

Review 2
--------

Review 3
--------

Review 4
--------

1. The reviewer notes that the work is not state of the art. I think it is
mentioned a couple of times (even in the abstract) what flowy can do that
other's cant and in "related work" it is listed how it has even more abilities
than SiLK.