### 2012-10-16 rebuttal.txt
some comments about stream-based

Review 1:
1. wrong track
2. lack of description of the language
3. too many implementation details
4. metrics used
5. research questions and conclusions

Review 2:
1. Gigascope, point to the original NFQL paper

Review 3:

- design advantages of the proposed tool are not explained, if this is
  because NFQL has already been proposed in another paper, then this
  work is derivative to a high degree
- again, because the proposed method (not tool) is not explained, and
  explanation of related tools lacks functional and methodological
  detail, it is difficult to judge WHY the proposed tool is better

The flow query language was introduced in [1]. It describes the _design_
advantages of the language. In fact, the expressiveness of the language
can be seen from [2] where we define queries to identify flow signatures
of popular applications. We have had a python prototype implementation
[3] in the past, however the prototype was extremely slow and could not
scale to real-world traces. This paper introduces NFQL, an efficient
implementation of the language. This is not just a reimplementation in
C, but the pipeline has been reimagined with crispier data structures to
allow the implementation to scale to real-world traces. This is the
first time we evaluate and show that NFQL is comparable to contemporary
flow-processing tools. This paper discusses these implementation aspects.

[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
Record Query Language,” DSOM ’09
[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
Popular Applications,” IM '11.
[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
stream-based IP flow record query language,” AIMS’10

- the method is a -QL -- a query language, which assumes fixed
  storage, while recent research on traffic processing moves away from
  the store-then-process approach

The query language of NFQL is _indeed_ stream-based. The processing
pipeline consists of a number of elements, which are connected with each
other via pipes. Each element gets a input _stream_, performs some sort
of operation on it and the output _stream_ is piped to the next element.
This only requires that NFQL maintains state information between the
elements. NFQL does not necessarily follow a store-then-process approach

# i do not think this is what the reviewer understands as "stream-based"
# nfql is not able to process packages as they come. Because it allows
# to do relative comparison in the grouper and the merger, those stages
# need to have access to any record of the dataset
# an extension that could make nfql stream-based for the grouper is a
# shifting window which only evaluates a small neighbourhood of the
# packet that is currently evaluated
# but this idea fails for the merger
#  - josch

Review 4:

- can you explain why you need the splitter before the filtering blocks
and not the other way round? IMHO filtering should happen before
splitting (fig 1)

The the order of execution of the pipeline stages in the flow query
language [1] and in the NFQL implementation is different.  According to
the language specification, as briefly described in section III:

"The pipeline starts with the splitter. It takes the flow-records data
as input in flow-tools compatible format. The splitter is responsible to
duplicate the input data out to several branches without any processing
whatsoever. This allows each of the branches to receive an identical
copy of the flow data to process it independently."

However, we (and as did the reviewer) realized that in order to
efficiently process large traces, the NFQL implementation needs to sway
away from this design decision. As such, NFQL is implemented in a way to
give away the need for a splitter entirely, as described in section IV-B

"NFQL uses identifiers to reference a flow record in the char array.
This eliminated the need to copy all the flow-records when moving ahead
in the processing pipeline. As a result, there is no dedicated splitter
stage in the execution engine. Each branch references the flow records
from a common memory location. This helps keep the memory costs at a
minimum when multiple branches are involved."

As a result, there is no splitter in the NFQL implementation.

[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
Record Query Language,” DSOM ’09.Berlin: Springer-Verlag, 2009, pp. 15–28

- can you explain why you do not filter flows on the fly?

We _do_ filter flows on the fly. This is described in section IV-B

"The execution engine, as defined by the flow query language must read
all the flow records of a supplied trace into memory before starting the
processing pipeline. Since, the filter stage uses the supplied set of
absolute rules to make a decision on whether or not to keep a flow
record; it has to pass through the whole in-memory recordset again to
produce the filter results. This technique involves multiple linear runs
on the trace and therefore slows down when the ratio of number of
filtered records to the total number of flow-records is high. We
optimized this behavior in NFQL by merging the filter stage with
in-memory read of the trace. This means, a decision on whether or not to
make room for a record in memory and eventually hold a pointer for it in
filter results is done upfront as soon as the record is read from the
trace. In addition, if a request to write the filter stage results to a
flow-tools file has been made, the writes are also made as soon the
filter stage decision is available, thereby allowing
reading-filtering-writing to happen in O(n) time, where n is the number
of records in the trace."

The previous prototype implementations (Flowy and Flowy 2.0) had this
issue. There was a separate filter stage and the flows were not filtered
on the fly. NFQL however _does_ filter flows on the fly. We even
evaluated the performance improvements of NFQL against Flowy 2.0 and saw
that NFQL now runs 10 times faster as shown in Fig 3.

- in IV.B you talk about indexes. What indexes have you used. Please cover
this topic better

Each flow record is stored as a char array. The trace is stored as an
array of these flow records. This allows NFQL to reference any flow
record using identifiers (indexes) in the array for O(1) retrieval.

- The work described on this paper is interesting but probably not the
state of the art in terms of performance and features (i.e. it's not a
step ahead but more a student project). Some examples are not easy to
understand.

We have tried to set forth the novelty of NFQL from the very outset,
The abstract for instance, clearly briefs out the functionality that
separates it from contemporary flow-processing tools:

"NFQL can process flow records, aggregate them into groups, apply
absolute or relative filters, invoke Allen interval algebra rules, and
merge group records."

In section II (related work), we reason how NFQL does more than the
surveyed contemporary flow-processing tools:

flow-tools and nfdump:
"The power of filtering rules in both the tools is however mostly
limited to absolute comparisons of flow attributes. As a result,
relative comparison amongst different flows or querying a timing
relationship among them is not possible." - which is possible in NFQL.
In fact, the capabilities of flow-tools and nfdump are just one of the
five stages of the NFQL's processing pipeline. They do not support
grouping, aggregation, merging et al.

SiLK: 
SiLK can group flows and merge group records. However, the parameters
over which grouping/merging can be done is faily limited:
"The grouping and merging operations can only be performed using
an equality operator. NFQL on the other hand, provides a much richer set
of comparisons operations, such as equal, not equal, greater than, less
than, greater or equal, less or equal."

"The newly formed groups, which are passed on to the group filter, can
also contain meta-information about the flow-records contained within
the group using the aggregate clause defined as part of the grouper
query. The possible aggregation operations are static, count, product,
sum, logical and/or/xor, arithmetic mean, standard deviation, union,
median, minimum and maximum." - whereas SiLK only supports sum and or
aggregation operations.

"The relative filtering on the groups are applied to express timing and
concurrency constraints using Allen interval algebra [6]." - The merger
can also match group records by putting timing characteristics under
considerations. No such Allen operations are possible in SiLK.

"The ungrouper is the final processing element of the pipeline. It
unwraps the tuples of group-records into individual flow-records,
ordered by their timestamps." - NFQL maintains flow membership information of
each group record. SiLK does not provide this feature. As a result, once
flows are grouped together in SiLK, there is no way to unfold the groups
to identify their member flows. As such, there is no ungrouper in SiLK.

In section V, we evaluate the performance of each tool against NFQL.
flow-tools and nfdump could only participate in one of the evaluations
due to their limited feature sets. In rest of the evaluations only SiLK
could participate. Even so, we had to carefully design queries to find a
common ground since most of the NFQL operations are not possible in SiLK.

We believe NFQL is major step ahead in this effort. Our first python
prototype Flowy was feature complete, but was extremely slow in
comparison to contemporary tools and therefore could not scale to
real-world traces.  NFQL is an efficient feature complete implementation
of our flow-query language. It is not just a reimplementation in C, but
the processing pipeline has been reimagined to allow it to scale to
real-world traces. This makes NFQL comparable to current flow-processing
tools, in addition to offering more feature-sets that are just not
available elsewhere.

It is not clear from the review which examples are not easy to
understand, so it's hard to comment.