### 2012-10-16 rebuttal.txt
* reducing rebuttal content   * currently at 4393 (should be below 3K)

Reviewer 1:
- main conclusions?

* NFQL, an efficient feature-complete implementation of the flow query
  language that supersedes our original python prototype Flowy [3].
* NFQL, can now process real-world sized traces in matter of minutes,
  something which used to take days (or even weeks) in Flowy. This
  expands the scope of flow-processing tools given the added feature set
  capability offered by the language which is not available elsewhere.
* The performance evaluation reveal that NFQL is comparable to
  contemporary flow-processing tools. SiLK appears faster in some situations.
* The evaluation queries developed as a part of this research can
  incubate into a more general benchmark suite for flow query tools

Reviewer 1, 3:
- flow query language description (design advantages)?
- example queries?

The flow query language was introduced in [1]. It describes the language
in full detail. This paper discusses the implementation aspects of the
language. We briefly describe the processing pipeline because that
helps create a stand ground necessary to understand the tool. Our
previous use-case paper [2] defines a number of queries that use the language.

Reviewer 1:
- 'Output Flows/Input Flows' metric?

The metric was used to blackbox each stage of the pipeline. This helped
benchmarks them separately to identify areas that may need attention.

Reviwer 1:
- used queries? reproduce/repeat your measurements?

The queries are in the git repository referenced as a footnote in the
paper. The JSON queries were omitted from the paper for brevity reasons
The measurements can be repeated by using the cited public trace
alongwith the queries in the repository. The benchmarking suite
developed as a part of the research will help automate the task.

Reviewer 4:
- filtering before splitting?

The filtering _does_ happen before splitting. In fact, NFQL is
implemented in a way to give away the need for a splitter entirely. As
such, there is no splitter in the NFQL implementation. This is described
in section IV-B: Splitter. This is one of the performance optimizations
that makes NFQL scale well with real-world traces.

Reviewer 4:
- filter flows on the fly?

We _do_ filter flows on the fly.
This is described in section IV-B: Filter

The previous prototype implementations (Flowy and Flowy 2.0) had this
issue. There was a separate filter stage and the flows were not filtered
on the fly. NFQL however _does_ filter flows on the fly. We even
evaluated the performance improvements of NFQL against Flowy 2.0 and saw
that NFQL now runs 10 times faster as shown in Fig 3.

Reviewer 2:
- older references (gigascope)?

The reference to Gigascope is in our flow query language paper [1],

Reviewer 2, 4:
- not state of the art?
- not a giant leap over prior work?

We have had a python prototype implementation [3] in the past, however
the prototype was extremely slow. NFQL is not just a reimplementation in
C, but the pipeline has been reimagined with better data structures to
allow the implementation to scale to real-world traces. We show that
NFQL is comparable to contemporary flow-processing tools. This paper
discusses these implementation aspects.

The abstract briefs out the novelty and is elaborated in further
sections. For instance, In section II, we reason how NFQL does more:

flow-tools and nfdump:
The capabilities of flow-tools and nfdump are just one of the
five stages of the NFQL's processing pipeline. They do not support
grouping, aggregation, merging et al.

SiLK: 
* SiLK can group flows and merge group records. However, the parameters
over which grouping/merging can be done is faily limited:
* SiLK only supports sum and or aggregation operations compared to a huge
set of aggregation options available in NFQL.
* The NFQL merger can also match group records by putting timing
characteristics under considerations. No such Allen operations are
possible in SiLK.
* Once flows are grouped together in SiLK, there is no way to unfold the
groups to identify their member flows. As such, there is no ungrouper in SiLK.

[1] V. Marinov and J. Schönwälder, “Design of a Stream-Based IP Flow
Record Query Language,” DSOM ’09
[2] V. Perelman, N. Melnikov, and J. Schönwälder, “Flow signatures of
Popular Applications,” IM '11.
[3] K. Kanev, N. Melnikov, and J. Schönwälder, “Implementation of a
stream-based IP flow record query language,” AIMS’10