### 2012-10-16 rebuttal.txt
rebuttal under 3000 characters

Reviewer 1:
- main conclusions?

* an efficient feature-complete implementation of the flow query
  language, supersedes our python prototype Flowy [3].
* can now process real-world sized traces in matter of minutes,
  something which used to take days in Flowy. This expands the scope of
  flow-processing tools, given the added feature set capability
  which is not available elsewhere.
* Performance evaluation reveal that NFQL is comparable to
  contemporary tools. SiLK appears faster in some situations.
* The evaluation queries developed can incubate into a more general 
  benchmark suite for flow query tools

Reviewer 1, 3:
- flow query language description (design advantages)?
- example queries?

The flow query language is described in full detail in [1]. This paper
discusses the implementation aspects of the language. We briefly
describe the processing pipeline because that helps create a stand
ground necessary to understand the tool. Our previous use-case paper [2]
defines a number of queries that use the language.

Reviewer 1:
- 'Output Flows/Input Flows' metric?

The metric was used to blackbox each stage of the pipeline. This helped
benchmarks them separately to identify areas that may need attention.

Reviwer 1:
- used queries? reproduce/repeat your measurements?

The queries are in the git repository referenced as a footnote in the
paper. The measurements can be repeated by using the cited public trace
alongwith the queries in the repository. The benchmarking suite
developed as a part of the research will help automate the task.

Reviewer 4:
- filtering before splitting?

The filtering _does_ happen before splitting. In fact, NFQL is
implemented in a way to give away the need for a splitter entirely. As
such, there is no splitter in the NFQL implementation. This is described
in section IV-B: Splitter.

Reviewer 4:
- filter flows on the fly?

We _do_ filter flows on the fly. This is described in section IV-B:
Filter Flowy and Flowy 2.0 had this issue. There was a separate filter
stage and the flows were not filtered on the fly. NFQL however _does_
filter flows on the fly. We evaluated the performance improvements
against Flowy 2.0 and saw that NFQL now runs 10 times faster as 
shown in Fig 3.

Reviewer 2:
- older references (gigascope)?

The reference to Gigascope is in our flow query language paper [1],

Reviewer 2, 4:
- not state of the art?
- not a giant leap over prior work?

NFQL is not just a reimplementation of Flowy in C, but the pipeline has
been reimagined with better data structures to allow NFQL to scale to
real-world traces. We show that NFQL is comparable to contemporary
flow-processing tools. This paper discusses these implementation aspects.

The abstract briefs out the novelty and is elaborated in further
sections. For instance, In section II, we reason how NFQL does more:

flow-tools and nfdump:
The capabilities of flow-tools and nfdump are just one of the
five stages of the NFQL's processing pipeline. They do not support
grouping, aggregation, merging et al.

SiLK: 
* SiLK can group flows and merge group records. However, the parameters
over which grouping/merging can be done is faily limited:
* SiLK only supports sum and or aggregation operations compared to a huge
set of aggregation options available in NFQL.
* The NFQL merger can also match group records by putting timing
characteristics under considerations. No such Allen operations are
possible in SiLK.
* Once flows are grouped together in SiLK, there is no way to unfold the
groups to identify their member flows. As such, there is no ungrouper in SiLK.

[1] “Design of a Stream-Based IP Flow Record Query Language,” DSOM ’09
[2] “Flow signatures of Popular Applications,” IM '11.
[3] “Implementation of a stream-based IP flow record query language,” AIMS’10