### 2012-10-16 rebuttal.txt
forgot to commit the amend

Reviewer 1:
- main conclusions?

* an efficient feature-complete implementation, 
  supersedes our python prototype Flowy [3].
* can now process real-world sized traces in minutes,
  This expands the scope of flow-processing tools, given the added 
  feature set capability which is not available in other tools.
* Evaluations reveal that NFQL is comparable to
  contemporary tools. SiLK appears faster in some situations.
* Evaluation queries can incubate into a more general 
  benchmark suite for flow query tools

Reviewer 1, 3:
- flow query language description (design advantages)?
- example queries?

The language is described in detail in [1]. This paper discusses the
implementation aspects. We briefly describe the pipeline to help create
a stand ground necessary to understand the tool. Our previous use-case
paper [2] defines a number of queries to identify application
flow-signatures.

Reviewer 1:
- 'Output Flows/Input Flows' metric?

The metric was used to blackbox each stage of the pipeline. This helped
benchmark them separately to identify areas that may need attention.

Reviwer 1:
- used queries? reproduce/repeat your measurements?

The queries are in the git repository referenced in a footnote in the
paper. The measurements can be repeated by using the cited public trace
alongwith the queries in the repository. The developed benchmarking suite
will help automate the task.

Reviewer 4:
- filtering before splitting?

The filtering _does_ happen before splitting. In fact, NFQL is
implemented in a way to give away the need for a splitter entirely. As
such, there is no splitter in the NFQL. This is described in section
IV-B: Splitter.

Reviewer 4:
- filter flows on the fly?

We _do_ filter flows on the fly. This is described in section
IV-B:Filter. Flowy and Flowy 2.0 had this issue. There is no longer a
separate filter stage in NFQL. NFQL now runs 10 times faster against
Flowy 2.0 as shown in Fig 3.

Reviewer 2:
- older references (gigascope)?

The reference to Gigascope is in our flow query language paper [1],

Reviewer 2, 4:
- not state of the art?
- not a giant leap over prior work?

NFQL is not just a reimplementation of Flowy in C, but the pipeline has
been reimagined with better data structures to allow NFQL to scale to
real-world traces. This makes NFQL comparable to current flow-processing
tools. This paper discusses these implementation aspects.

The novelty of NFQL is discussed in section II:

* flow-tools/nfdump features are just one of the 5 stages of the pipeline. 
* SiLK only supports grouping/merging on an equality operator
* SiLK only supports sum and or aggregation operations. 
* SiLK does not support Allen operations. 
* SiLK cannot unfold groups to identify their member flows.

[1] “Design of a Stream-Based IP Flow Record Query Language,” DSOM ’09
[2] “Flow signatures of Popular Applications,” IM '11.
[3] “Implementation of a stream-based IP flow record query language,” AIMS’10