### 2013-01-14 sections/implementation.tex

The \ac{NFQL} architecture primarily consists of a front-end parser backed up
by an execution engine as shown in Fig. \ref{fig:nfql-pipeline}. The execution
engine is the brain of \ac{NFQL} where the complete pipeline is processed. It
receives the flow query at runtime using a JSON intermediate format. The
execution engine is written in C, however the front-end parser can be written
in any desired language.

%\begin{figure}[ht!]
  %\begin{center}
    %\includegraphics* [width=0.8\linewidth]{nfql-architecture}
    %\caption{NFQL Architecture}
    %\label{fig:architecture}
  %\end{center}
%\end{figure}

\subsection{Flow Query Intermediate Format}
\input{sections/implementation/flow-query-intermediate-format}
\label{subsec:intermediate-format}

\subsection{Execution Workflow}
\input{sections/implementation/execution-workflow}
\label{subsec:execution-workflow}

\subsection{Performance Optimizations}
\input{sections/implementation/performance-optimizations}
\label{subsec:performance-optimizations}

\subsection{Adaptable Compression Levels}
\input{sections/implementation/adaptable-compression-levels}
\label{subsec:adaptable-compression-levels}