### 2013-01-14 sections/implementation.tex
changed NFQL tool to \texttt{nfql}

The \texttt{nfql} architecture primarily consists of a front-end parser backed
up by an execution engine as shown in Fig. \ref{fig:nfql-pipeline}. The
execution engine is the brain of \texttt{nfql} where the complete pipeline is
processed. It receives the flow query at runtime using a JSON \cite{rfc4627}
intermediate format. The execution engine is written in C, however the
front-end parser can be written in any desired language.

\subsection{Flow Query Intermediate Format}
\input{sections/implementation/flow-query-intermediate-format}
\label{subsec:intermediate-format}

\subsection{Execution Workflow}
\input{sections/implementation/execution-workflow}
\label{subsec:execution-workflow}

\subsection{Performance Optimizations}
\input{sections/implementation/performance-optimizations}
\label{subsec:performance-optimizations}

\subsection{Adaptable Compression Levels}
\input{sections/implementation/adaptable-compression-levels}
\label{subsec:adaptable-compression-levels}